Light inextensible strings under tension in the Schwarzschild geometry

### 2019-02-19 string_ajp.tex
rewording and add Abramowicz as a ref

\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim T/\left(\mu g\right)$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole gives $L\lesssim TGM/\left(\mu c^2\right)$.  For the string
to behave as though light on lengthscales comparable with the
Schwarzschild radius would require $T\gtrsim\mu c^2$.  Exceeding this
criterion would violate the dominant energy condition (and in this
context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
light inextensible string is an interesting and informative limit: its
analysis interprets the concept ``straight line'' and suggests a
possible way of manipulating a black hole.  The Schwarzschild metric

@@

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the left, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced, and the string is
inextensible so stores no energy.

>>>>

\label{intro}
Light inextensible strings are a stalwart feature of elementary
physics.  A light string is not subject to gravity, and being under
tension means that its configuration is an extremum of proper length.
Of course, no real string is perfectly weightless or inextensible.
With tension~$T$, gravitational acceleration $g$ and (proper) mass per
(proper) unit length~$\mu$, then a string will behave as though light
on lengthscales $L\lesssim T/\left(\mu g\right)$.  Using Newtonian
expressions for the strength of gravity at the event horizon of a
black hole gives $L\lesssim TGM/\left(\mu c^2\right)$.  For the string
to behave as though light on lengthscales comparable with the
Schwarzschild radius would require $T\gtrsim\mu c^2$.  Exceeding this
criterion would violate the dominant energy condition (and in this
context we note that transverse waves would propagate at
speed~$\sqrt{T/\mu}\sim c$).  Nevertheless, the idealization of a
string as light and inextensible is an interesting and informative
limit: its analysis interprets the concept ``straight line'' and
suggests a possible way of manipulating a black hole.  Peculiar
counterintuitive behaviour such as inward-directed centrifugal force
in the vicinity of a black hole is well
known~\citep{abramowicz1992,abramowicz1993}, and the analysis of taut
strings furnishes another everyday window into general relativity.
Note in particular that Abramowicz states ``feels with his hands''

The Schwarzschild metric

@@

In Figure~\ref{closest_approach_non_self_intersecting}, one's
classical intuition would suggest that the black hole feels a force to
the left, and should start accelerating.  However, it is not clear by
what mechanism such a force would be transmitted: the situation is
steady, so no gravitational waves are produced, and the string is
inextensible so stores no energy.  Further work might include analysis
of taut strings near a spinning black hole.
